### 2019-12-06 si.tex
Update si authors.

\documentclass[10pt,a4]{article}
\usepackage[top=0.85in,left=.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% multirow for multiple row tables. 
\usepackage{multirow}

\usepackage{subfig}

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\DeclareGraphicsExtensions{.pdf,.png}   
\usepackage{epstopdf}

\renewcommand\thefigure{S\arabic{figure}} 
\renewcommand\thetable{S\arabic{table}} 
%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Supplementary Material: Joint modelling of aggregated incidence data and point prevalence surveys for malaria mapping} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Tim C.D. Lucas*\textsuperscript{1}, 
Anita K. Nandi\textsuperscript{1},
Elisabeth G. Chestnutt\textsuperscript{1},
Katherine A. Twohig\textsuperscript{1},
Suzanne H. Keddie\textsuperscript{1},
Emma L. Collins\textsuperscript{1},
Rosalind E. Howes\textsuperscript{1},
Michele Nguyen\textsuperscript{1},
Susan F. Rumisha\textsuperscript{1},
Andre Python\textsuperscript{1},
Rohan Arambepola\textsuperscript{1},
Amelia Bertozzi-Villa\textsuperscript{1,2},
Penelope Hancock\textsuperscript{1},
Punam Amratia\textsuperscript{1},
Katherine E. Battle\textsuperscript{1},
Ewan Cameron\textsuperscript{1},
Peter W. Gething\textsuperscript{1, 3, 4}
Daniel J. Weiss\textsuperscript{1}
\\
\bigskip
1. BDI, Oxford\\
2. Institute for Disease Modeling, Bellevue, WA, USA\\
3. Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia\\
4. Curtin University, Perth, Australia
\\
\bigskip

\end{flushleft}

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%

% Current address notes

\section{Crossvalidation schemes}

\begin{figure}[h!]

\centering

\includegraphics[width = 0.8\textwidth]{figures/random_crossvalidation_full.png}

\caption{Random cross-validation scheme for Indonesia, Senegal and Madagascar. The fold for both aggregated incidence data and prevalence point data is shown.}
\label{fig:cv_random}
\end{figure}

\begin{figure}[h!]
\centering

\includegraphics[width = 0.8\textwidth]{figures/spatial_crossvalidation_full.png}

\caption{ Spatial cross-validation scheme for Indonesia, Senegal and Madagascar. The fold for both aggregated incidence data and prevalence point data is shown.}
\label{fig:cv_spatial}
\end{figure}

\clearpage
\section{Estimated parameters}

% latex table generated in R 3.4.4 by xtable 1.8-4 package
% Fri Jul 19 12:48:02 2019
\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
Parameters & Baseline & BaselineSD & PrevalenceGP & PrevalenceGPSD & Joint & JointSD \\ 
  \hline
$\beta_0$ & -5.58 & 0.29 & -4.66 & 0.58 & -5.35 & 0.52 \\ 
  $\beta_p$ &  &  &  &  & 1.63 & 0.13 \\ 
  LST day mean & 0.00 & 0.26 & -0.02 & 0.26 & -0.04 & 0.24 \\ 
  EVI & 0.43 & 0.23 & 0.35 & 0.24 & 0.35 & 0.20 \\ 
  TSI & -0.61 & 0.24 & -0.51 & 0.24 & 0.42 & 0.17 \\ 
  accessability & -0.09 & 0.21 & -0.07 & 0.21 & 0.11 & 0.11 \\ 
  elevation & 0.44 & 0.21 & 0.36 & 0.22 & 0.31 & 0.18 \\ 
  LST day SD & 0.23 & 0.20 & 0.07 & 0.21 & 0.24 & 0.20 \\ 
  Night lights & 0.10 & 0.06 & 0.08 & 0.06 & 0.07 & 0.04 \\ 
  TCW & -0.02 & 0.28 & -0.05 & 0.28 & 0.17 & 0.23 \\ 
  Prev GP &  &  & 0.30 & 0.17 &  &  \\ 
  $\omega_v$ & 0.25 & 0.25 & 0.27 & 0.25 & -0.03 & 0.32 \\ 
  $\omega_w$ &  &  &  &  & -1.03 & 0.13 \\ 
  $\log(\sigma_u)$ & -2.48 & 0.61 & -2.53 & 0.63 & -3.35 & 0.52 \\ 
  $\log(\rho)$ & 0.58 & 0.30 & 0.58 & 0.30 & 0.71 & 0.31 \\ 
  $\log(\alpha)$ &  &  &  &  & -0.00 & 0.00 \\ 
   \hline
\end{tabular}
\label{parssen}
\caption[Parameter estimates for Senegal]{Parameter estimates for Senegal. $\omega_v = \log\left(\frac{1}{{\sigma_v}^2}\right)$ and $\omega_w = \log\left(\frac{1}{{\sigma_w}^2}\right)$} 
\end{table}

% latex table generated in R 3.4.4 by xtable 1.8-4 package
% Fri Jul 19 12:48:37 2019
\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
Parameters & Baseline & BaselineSD & PrevalenceGP & PrevalenceGPSD & Joint & JointSD \\ 
  \hline
$\beta_0$ & -3.81 & 0.32 & -2.66 & 0.29 & -3.77 & 0.30 \\ 
  $\beta_p$ &  &  &  &  & 0.07 & 0.10 \\ 
  LST day mean & 0.16 & 0.29 & 0.14 & 0.27 & 0.14 & 0.24 \\ 
  EVI & 0.34 & 0.21 & 0.23 & 0.19 & 0.46 & 0.13 \\ 
  TSI & 0.45 & 0.22 & 0.42 & 0.21 & 0.63 & 0.19 \\ 
  accessability & 0.36 & 0.11 & 0.15 & 0.10 & 0.24 & 0.07 \\ 
  elevation & -0.11 & 0.18 & -0.05 & 0.16 & 0.31 & 0.12 \\ 
  LST day SD & -0.22 & 0.21 & -0.15 & 0.17 & -0.20 & 0.15 \\ 
  Night lights & -0.01 & 0.03 & -0.02 & 0.02 & -0.02 & 0.02 \\ 
  TCW & -0.26 & 0.26 & -0.22 & 0.25 & -0.15 & 0.15 \\ 
  Prev GP &  &  & 0.36 & 0.07 &  &  \\ 
  $\omega_v$ & 1.19 & 0.23 & 1.26 & 0.23 & 1.38 & 0.26 \\ 
  $\omega_w$ &  &  &  &  & -0.77 & 0.11 \\ 
  $\log(\sigma_u)$ & -2.57 & 0.60 & -2.83 & 0.63 & -3.21 & 0.49 \\ 
  $\log(\rho)$ & 0.99 & 0.32 & 0.69 & 0.31 & 0.73 & 0.27 \\ 
  $\log(\alpha)$ &  &  &  &  & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\label{parsmdg}
\caption{Parameter estimates for Madagascar $\omega_v = \log\left(\frac{1}{{\sigma_v}^2}\right)$ and $\omega_w = \log\left(\frac{1}{{\sigma_w}^2}\right)$} 
\end{table}

% latex table generated in R 3.4.4 by xtable 1.8-4 package
% Fri Jul 19 12:48:41 2019
\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
Parameters & Baseline & BaselineSD & PrevalenceGP & PrevalenceGPSD & Joint & JointSD \\ 
  \hline
$\beta_0$ & -5.83 & 0.65 & -5.59 & 0.80 & -6.31 & 0.56 \\ 
  $\beta_p$ &  &  &  &  & 0.03 & 0.20 \\ 
  LST day mean & -0.03 & 0.20 & -0.03 & 0.20 & 0.28 & 0.27 \\ 
  EVI & 0.11 & 0.08 & 0.11 & 0.08 & 0.55 & 0.18 \\ 
  TSI & 0.24 & 0.19 & 0.24 & 0.19 & -0.02 & 0.24 \\ 
  accessability & 0.73 & 0.24 & 0.73 & 0.24 & 0.36 & 0.19 \\ 
  elevation & -0.12 & 0.12 & -0.12 & 0.12 & -0.23 & 0.22 \\ 
  LST day SD & 0.04 & 0.11 & 0.04 & 0.11 & -0.04 & 0.18 \\ 
  Night lights & -0.34 & 0.12 & -0.34 & 0.12 & 0.01 & 0.10 \\ 
  TCW & 0.05 & 0.10 & 0.05 & 0.10 & 0.06 & 0.19 \\ 
  Prev GP &  &  & 0.06 & 0.12 &  &  \\ 
  $\omega_v$ & -1.54 & 0.11 & -1.54 & 0.11 & -1.33 & 0.11 \\ 
  $\omega_w$ &  &  &  &  & -2.65 & 0.09 \\ 
  $\log(\sigma_u)$ & -2.23 & 0.43 & -2.25 & 0.43 & -2.79 & 0.48 \\ 
  $\log(\rho)$ & 1.63 & 0.23 & 1.62 & 0.23 & 1.38 & 0.24 \\ 
  $\log(\alpha)$ &  &  &  &  & -0.00 & 0.00 \\ 
   \hline
\end{tabular}
\label{parsidn}
\caption{Parameter estimates for Indonesia $\omega_v = \log\left(\frac{1}{{\sigma_v}^2}\right)$ and $\omega_w = \log\left(\frac{1}{{\sigma_w}^2}\right)$} 
\end{table}

\clearpage
\section{Cross-validation predictions}

\begin{figure}[h!]
% to be removed before submission
\makebox{\includegraphics[width = 0.7\textwidth]{figures/idn_both_cv12_random_preds.png}}
\caption{\label{predobsmapsen1}
Incidence data and predicted incidence maps.
The incidence (log10) data (top), predicted log10 incidence from the prevalence GP model for random cross-validated out-of-sample polygons (middle) and predicted log10 incidence from the joint model for spatially cross-validated out-of-sample polygons (bottom) for Senegal.
}
\end{figure}

\begin{figure}[h!]
% to be removed before submission
\makebox{\includegraphics[width = 0.9\textwidth]{figures/mdg_both_cv12_random_preds.png}}
\caption{\label{predobsmapsen2}
Incidence data and predicted incidence maps.
The incidence (log10) data (top), predicted log10 incidence from the prevalence GP model for random cross-validated out-of-sample polygons (middle) and predicted log10 incidence from the joint model for spatially cross-validated out-of-sample polygons (bottom) for Senegal.
}
\end{figure}

\begin{figure}[h!]
% to be removed before submission
\makebox{\includegraphics[width = 1\textwidth]{figures/sen_both_cv12_random_preds.png}}
\caption{\label{predobsmapsen3}
Incidence data and predicted incidence maps.
The incidence (log10) data (top), predicted log10 incidence from the prevalence GP model for random cross-validated out-of-sample polygons (middle) and predicted log10 incidence from the joint model for spatially cross-validated out-of-sample polygons (bottom) for Senegal.
}
\end{figure}

\clearpage
\subsection{Full model predictions}
\subsubsection{Madagascar}

\begin{figure}[h!]
\centering

\includegraphics[width = 0.3\textheight]{figures/mdg_baseline_api.png}

\caption{Predictions from the baseline model fitted to all data in Madagascar.}
\label{baselinemdg}
\end{figure}
\begin{figure}[h!]
\centering

\includegraphics[width = 0.3\textheight]{figures/mdg_prgp_api.png}

\caption{Predictions from the prevalence GP model fitted to all data in Madagascar.}
\label{gpmdg}
\end{figure}

\begin{figure}[h!]
\centering

\includegraphics[width = 0.3\textheight]{figures/mdg_joint_api.png}

\caption{Predictions from the joint model fitted to all data in Madagascar.}
\label{jointmdg}
\end{figure}

\subsubsection{Senegal}

\begin{figure}[h!]
\centering

\includegraphics[width = 0.3\textheight]{figures/sen_baseline_api.png}

\caption{Predictions from the baseline model fitted to all data in Senegal.}
\label{baselinesen}
\end{figure}
\begin{figure}[h!]
\centering

\includegraphics[width = 0.3\textheight]{figures/sen_prgp_api.png}

\caption{Predictions from the prevalence GP model fitted to all data in Senegal.}
\label{gpsen}
\end{figure}

\begin{figure}[h!]
\centering

\includegraphics[width = 0.3\textheight]{figures/sen_joint_api.png}

\caption{Predictions from the joint model fitted to all data in Senegal.}
\label{jointsen}
\end{figure}

\subsubsection{Indonesia}

\begin{figure}[h!]
\centering

\includegraphics[width = 0.5\textheight]{figures/idn_baseline_api.png}

\caption{Predictions from the baseline model fitted to all data in Indonesia.}
\label{baselineidn}
\end{figure}
\begin{figure}[h!]
\centering

\includegraphics[width = 0.5\textheight]{figures/idn_prgp_api.png}

\caption{Predictions from the prevalence GP model fitted to all data in Indonesia.}
\label{gpidn}
\end{figure}

\begin{figure}[h!]
\centering

\includegraphics[width = 0.5\textheight]{figures/idn_joint_api.png}

\caption{Predictions from the joint model fitted to all data in Indonesia.}
\label{jointidn}
\end{figure}

\clearpage

\subsection{Full model fields}

\subsubsection{Senegal}

\begin{figure}[h!]
\centering

\includegraphics[width =  0.3\textheight]{figures/sen_baseline_field.png}

\caption{The random field as fitted to all the data in the baseline model.}
\label{baselinefieldsen}
\end{figure}

\begin{figure}[h!]
     \centering
     \subfloat[][The prevalence GP model that is used as a covariate.]{\includegraphics[width = 0.3\textheight]{figures/sen_prevgp.png}}
     \subfloat[][The random field as fitted to all the data in the prevalence GP model.]{\includegraphics[width = 0.3\textheight]{figures/sen_prgp_field.png}}
     \label{gpsencov}
\end{figure}
%remaining part

\begin{figure}[h!]
\centering

\includegraphics[width =  0.3\textheight]{figures/sen_joint_field.png}

\caption{The random field as fitted to all the data in the joint model.}
\label{jointfieldsen}
\end{figure}

\clearpage
\subsubsection{Indonesia}

\begin{figure}[h!]
\centering

\includegraphics[width =  0.6\textheight]{figures/idn_baseline_field.png}

\caption{The random field as fitted to all the data in the baseline model.}
\label{baselinefieldidn}
\end{figure}

\begin{figure}[h!]
     \centering
     \subfloat[][The prevalence GP model that is used as a covariate.]{\includegraphics[width =0.5\textwidth ]{figures/idn_prevgp.png}}
     \subfloat[][The random field as fitted to all the data in the prevalence GP model.]{\includegraphics[width = 0.5\textwidth]{figures/idn_prgp_field.png}}
     \label{gpsencov}
\end{figure}
%remaining part

\begin{figure}[h!]
\centering

\includegraphics[width =  0.6\textheight]{figures/idn_joint_field.png}

\caption{The random field as fitted to all the data in the joint model.}
\label{jointfieldidn}
\end{figure}

\clearpage
\subsubsection{Madagascar}

\begin{figure}[h!]
\centering

\includegraphics[width =  0.3\textheight]{figures/mdg_baseline_field.png}

\caption{The random field as fitted to all the data in the baseline model.}
\label{baselinefieldmdg}
\end{figure}

\begin{figure}[h!]
     \centering
     \subfloat[][The prevalence GP model that is used as a covariate.]{\includegraphics[width =0.5\textwidth ]{figures/mdg_prevgp.png}}
     \subfloat[][The random field as fitted to all the data in the prevalence GP model.]{\includegraphics[width = 0.5\textwidth]{figures/mdg_prgp_field.png}}
     \label{gpsencov}
\end{figure}
%remaining part

\begin{figure}[h!]
\centering

\includegraphics[width =  0.3\textheight]{figures/mdg_joint_field.png}

\caption{The random field as fitted to all the data in the joint model.}
\label{jointfieldidn}
\end{figure}

\end{document}